Exception Handling Bug Hazards in Android

### 2015-10-11 SPRINGER_JOURNAL/msr_journal.tex
fixing typos
grammar and spelling

\begin{abstract}
This paper reports two studies:  a repository mining study whose goal was to identify 
exception handling bug hazards in Android apps; and a survey-based study whose goal
was to assess the perspectives of developers concerning such bug hazards.
The repository mining study mined the exception stack traces embedded on 
159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
and Googlecode (157 projs.).  Overall 6,005 exception stack traces were extracted, 
and analyzed in combination with source code and bytecode analysis. 
Some bug hazards were consistently detected such as:  unexpected exception wrappings 
(e.g., Errors being wrapped in checked exceptions), and NullPointerExeptions 
as the major cause of reported crashes - reported in 
28\% of all mined stack traces found in about 50\% of all analyzed projects. 
Finally, a survey was applied to some developers 
involved on some of the GitHub Android projects analyzed in our study 
and their perspective about the bug hazards and 
how they deal with them was assessed and discussed.

@@

The number of mobile apps is increasing in a daily rate. If on the one hand they
extend phones capabilities far beyond of the basic calls, on the other hand they have to 
cope with an increasing number of exceptional conditions 
(e.g., faults in underlying middleware or hardware; compatibility issues~\cite{McDon13}; 
memory and battery restrictions; noisy external resources~\cite{Zhang12}).

@@

However, such mechanisms are more often then not the less 
understood and tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling,yuan:2014.osdi}.
As a consequence they may inadvertently negatively affect the system: exception-related code may introduce failures such as 
uncaught exceptions~\cite{jo2004uncaught, Zhang12} - 
which can lead to system crashes, making the system even less robust~\cite{coelho2011unveiling}.

In Java, when an application fails due to an uncaught exception, 
it automatically terminates, while the system prints a stack trace to the console, 
or on a log file~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
of the thrown exception and the ordered list of methods that were active on the call stack before 
the exception occurred~\cite{gosling2000java,bloch2008effective}.
When available, the exception stack traces provide a useful source of information about system crashes ~\cite{bettenburg2008makes} which 
can enable different kinds of post-mortem analysis and support:  debugging~\cite{schroter2010stack}, 
bug classification and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  
automated bug fixing~\cite{sinha2009fault} and fault-proneness prediction models~\cite{kim2013predicting}.

@@

This work is conducted in two phases. Firstly, a mining study performs a post 
mortem analysis of the exception stack traces 
included in issues reported on Android projects hosted in GitHub and Google Code.
The goal of this study is to investigate whether the reported exception stack traces
 can reveal common \emph{bug hazards} in the exception-related code.  
A \emph{bug hazard}~\cite{binder2000testing}  is a circumstance that increases the 
chance of a bug being present in the software. An example of a bug hazard can 
be a characteristic of  the exception-related code which can increase the likelihood 
of introducing the aforementioned uncaught exceptions.
Secondly, an exploratory survey to the Android developers involved on the mined projects
to assess their perspective about the exception handling bug hazards found.

To guide this investigation we compiled general guidelines on how to use Java
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.
Then, using a custom tool called ExceptionMiner,
 which we developed specifically for this study, we mine stack traces from the issues reported 
on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
The exception stack trace analysis was augmented by means of bytecode 
and source code analysis on the exception-related code of the Android 
platform and Android applications. Some \emph{bug hazards} consistently detected 
during this mining study include:

@@

   \item  Cross-type exception wrappings, such as an OutOfMemoryError wrapped in a checked exception.
Trying to handle an instance of OutOfMemoryError ``hidden" in  a checked exception may bring the program
 to an unpredictable state. Such wrappings suggest that when (mis)applied, the exception wrapping can make 
the exception-related code more complex and negatively impact the application robustness.

  \item  Undocumented runtime exceptions raised by the Android platform (35 methods) and  third-party libraries (44 methods) -
 which correspond to 4.4\% of the reported exception stack traces.
In the absence of the ``exception specification" of third-party code, it is difficult or 
even infeasible for the developer to protect the code against ``unforeseen'' exceptions. 
Since in such cases the client usually does not have access to the source code, such undocumented 
exceptions may remain uncaught and lead to system crashes.

@@

 \item A multitude of programming mistakes - approximately 52\% of the reported stack traces can 
be attributed to programming mistakes. In particular, 27.71\% of all stack traces contained a java.lang.NullPointerException 
as their root cause.

@@

The exploratory survey was conducted with 70 Android developers involved in one or more of the 
GitHub Android projects whose issues were mined. This survey reveals that only few developers (3\% of respondents)
 new about the undocumented checked exceptions signaled by native C code of Android platform. 
Moreover, most of the developers recognized that cross-type exception wrappings may
negatively impact the application robustness. The uncaught exceptions due to errors in programming 
logic, with the NullPointerException was identified by most developers (68\%) as the first or second main cause 
of application crashes.

 The high prevalence of NullPointerExceptions found in the mining study and confirmed on the exploratory survey
 is in line with findings of earlier research~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}, 
as are the undocumented runtime exceptions signaled by the Android Platform~\cite{kechagia2014}.

Some of the findings of our mining study emphasize the impact of these bug hazards on the application robustness 
by mining a different source of information as the ones used in previous works. The present work 
mined issues created by developers on GitHub and Google Code, while  previous research analyzed crash reports and automated test reports.
 Furthermore, our work points to bug hazards that were not detected by previous research (i.e., cross-type wrappings, undocumented checked exceptions and undocumented runtime exceptions thrown by third-party libraries) which represent new threats to application robustness.
Moreover, we perform  the first exploratory survey study whose
goal was to assess developers perspective regarding a set of exception handling bug hazards
as well as how developers deal with exceptions and prevent crashes while developing.

Our findings point to threats not only to the development of robust Android apps, 
but also to the development of any robust Java-based system. 
Hence, the study results are relevant to Android and Java developers who may underestimate the effect of such
 \emph{bug hazards} on the application robustness,
and who have to face the difficulty of preventing them.
Moreover, such \emph{bug hazards} 
call for improvements on languages (e.g. to prevent null pointer dereferences) and tools to better support exception handling in Android and Java environments.

The remainder of this paper is organized as follows. 
Section~\ref{sec:back} provides the necessary
background on the Android platform and the Java exception model. 
Section~\ref{sec:study} presents the study design. 
Section~\ref{sec:exceptionminer} describes the ExceptionMiner tool we developed to conduct our study.
Section~\ref{sec:result} reports the findings of our study.
Section~\ref{sec:disc} provides a discussion of the wider implications of our results,
 presents the threats to validity associated to the mining study and discusses the limitations
of the survey-based study. 
Finally Section~\ref{sec:rele} describes related work, 
and Section~\ref{sec:conc} concludes the paper and outlines
directions for future work.
\section{Background}
\label{sec:back}

@@

\emph{\textbf{Exception Types.}} In Java, exceptions are represented according to a class hierarchy, on which
 every exception is an instance of the Throwable class, and can be of three kinds: the checked exceptions
(extends Exception), the runtime exceptions (extends RuntimeException) and errors
(extends Error)~\cite{gosling2000java}. Checked exception received their name
 because they must be declared on the method's \emph{exception interface} (i.e., the list of exceptions that a method 
might raise during its execution) and the compiler statically checks if
 appropriate handlers are provided within the system.
Both runtime exceptions and errors are also known as ``unchecked exceptions'', as 
they do not need to be specified on the method \emph{exception interface} and do not trigger any 
compile time checking.

By convention, instances of Error represent  unrecoverable conditions which usually result
from failures detected by the Java Virtual Machine due to resource limitations, such as OutOfMemoryError.
Normally these cannot be handled inside the application.  Instances of RuntimeException are implicitly 
thrown by Java runtime environment when a program violates 
the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
error, null pointer references). Some programming languages react to such errors by immediately terminating the program, while
other languages, such as C++, let the program  continue
 its execution in some situations such as the out-of-bounds array index.
According to the Java Specification~\cite{gosling2000java} programs are not 
expected to handle such runtime exceptions signaled by the runtime environment.

User-defined exceptions can be either checked
 or unchecked, by extending either Exception or RuntimeException. There is a long-lasting debate 
about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
%\footnote{152 questions in Stackoverflow are related to this debate}. 
Section~\ref{sec:best} presents
a set of best practices related to each of them.

@@

 A common way of  propagating exceptions in Java programs is through
 exception wrapping (also called chaining):
One exception 
is caught and wrapped in another which is then thrown instead. Figure~\ref{fig:wrapping} shows 
an exception stack trace which illustrates such exception wrapping. 
For simplicity, in this paper we will refer to ``exception stack trace'' as as just stack trace.
The bottom part of the stack trace is the \emph{root exception} (Figure~\ref{fig:wrapping}-A), which indicates
the first reason (root cause) for the exception thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation, which we will refer to as the \emph{exception wrapper} in this paper (Figure~\ref{fig:wrapping}-C). The
execution flow  between the root exception and the wrapper may
include other intermediate exception wrappers (Figure~\ref{fig:wrapping}-D). In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration (Figure~\ref{fig:wrapping}-B).

@@

Our study focused on the exception stack traces contained in issues of Android projects (hosted in GitHub and Google Code).
And to support our investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
which extracts the exception stack traces embedded on issues, 
and combines stack trace information with source code and bytecode 
analysis. Moreover, we use manual inspection to augment
 the understanding of stack traces and support further discussions and insights (Section~\ref{sec:manual}).
In this study we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively.

Figure~\ref{fig:overview} gives an overview of our study.
First,  the issues reported on Android projects hosted on GitHub (1) and Google Code (2) are recovered.
Then the stack traces embedded in each issue are extracted and distilled (3).
The stack trace information is then combined with source code and bytecode analysis in order to discover 
the type of the exceptions (5) reported on the stack traces (e.g., error, runtime, checked),
and the origin (6) of such exceptions 
(e.g., the application, a library, the Android platform). 
Manual inspection steps (4, 7, 9) are used to support the mining process  and 
the search for \emph{bug hazards}  (8). The next sections detail each step of this mining process.

Our study focuses on open-source apps, since the information needed to perform our study
cannot be retrieved from commercial apps, whose issue report systems and 
source codes are generally not publicly available. 
Open source Android open-source apps have also been the 
target of other research~\cite{Linar13,Ruiz12} addressing \emph{reuse} and API stability.

@@

\subsection{Android Apps in GitHub}
\label{sec:git}

@@

This study uses the dataset provided by the GHTorrent project~\cite{Gousi13}, 
an off-line mirror of the data  offered through the Github API.  
To identify Android projects, we performed a case insensitive search for the
term ``android" in the repository's names and short descriptions.  
Up to 23 February 2014,  when we queried GHTorrent, this resulted in 2,542 repositories. Running the ExceptionMiner tool 
 in this set we observed that 589 projects had at least one issue containing a stack trace.

Then we performed a further clean up, inspecting the site of every Android project
reporting at least one stack trace, to make sure that they represented real
mobile apps. During this clean up 107 apps were removed because they were either
example projects (i.e., toy projects) or tools to support Android development
(e.g. Selendroid, Roboeletric - tools to support the testing of Android apps).
The filtered set consisted of 482 apps. This set of 482 projects contained overall 31,592 issues from which 4,042 exception stack traces 
were extracted.

Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none or an arbitrary set of labels depending 
on its repository. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
on the issues including exception stack traces. Regardless of the issue labels, every exception stack
trace may contain relevant information concerning the exception structure of the
projects analyzed, and therefore can reveal \emph{bug hazards} on the exception-related code.
Because of this, we opted for not restricting the analysis to just defect issues.

@@

\subsection{Android Apps in Google Code}
Google Code contains widely used open-source Android apps (e.g. K9Mail\footnote{K9Mail moved to Github but as a way of not loosing the project history it advises their users to report bugs on the Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}).
However, differently from GitHub, Google Code does not provide an API to access the information related
 to hosted projects.\footnote{Google code used to provide a Web service to its repositories, but this was deactivated in June 2013 in what Google called a ``clean-up action''.}
To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigates 
through the web interface of Google Code projects extracting all issues and issue comments and storing them in a relational database for later analysis.
To identify Android projects in Google Code, we performed a similar heuristic: we performed a case insensitive search 
(on the Google Code search interface) for the term ``android''. On January 2014, when we queried Google Code, this resulted in a list of 788  projects. This list comprised the seeds sent to our Crawler.

The Crawler retrieved all issues and comments for these projects.
From this set, 724 projects defined at least 1 issue. Running the ExceptionMiner tool 
 in this set we observed that 183 projects had at least one issue containing an exception stack trace.
 Then we performed further clean up (similar to the one described previously) inspecting the site 
of each project. As a result we could identify 157 Android projects.  This set contained  127,456 issues in total,
 from which 1,963 exception stack traces were extracted. Table~\ref{tab:lables} illustrates the occurrences of different labels 
on the issues including exception stack traces. Differently from GitHub, in Google Code most of 
the issues were labeled as ``Defect''. However, based on the same assumption described for the GitHub repository
 we considered all issues reporting stack traces (regardless its labels).

@@

\emph{\textbf{Repository Connectors.}} This component enables the connection 
with issue repositories. In this study two main connectors were created: one which connects to 
GHTorrent database, and a Google Code connector which is comprised of a Web Crawler
 that can traverse the Google Code web interface and extract the project's issues. 
Project meta-data and the issues associated with each project are stored in a relational 
database.

\emph{\textbf{Exception Stack Trace Distiller.}}
This component combines a parser (based on regular expressions) 
and heuristics able to identify and filter exception names and stack traces inline with text. 
This component distills the information that composes a stack trace.
 Some of the attributes extracted from the stack trace are
 the root exception and its signaler, as well as the exception wrappers and their corresponding signalers. 
This component also distills fine grained information of each attribute such as the classes and packages associated to them.
In contrast to existing issue parsing solutions such as Infozilla, our parser
 can discover stack traces mixed with log file information\footnote{In several 
exception stack traces, the exception frames were preceeded by logging information e.g., 
\texttt{03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread\-.java:127)} which could not be detected by existing tools.}.

\emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
stack traces every exception defined on a stack trace needs to be classified according to its type
(e.g. Error, checked Exception or RuntimeException). The module responsible for this analysis 
uses the Design Wizard framework \cite{Brunet09} to inspect the bytecode of the exceptions reported 
on stack traces.  It walks up the type hierarchy of a Java exception until it reaches 
a base exception type. Hence in this study the bytecode analysis was used to discover the type 
of each mined exception when the jar file of such exception was available on the project or 
on a reused  Java library. A specific implementation (based on source code analysis)
 was needed to discover the exception type when the bytecode was not available.
With this module we analyzed all exceptions defined in the Android platform (Version 4.4, API level 19),
which includes all basic Java exceptions that can be thrown during the app execution,
and exceptions thrown by Android core libraries. Moreover, we also analyzed the 
exceptions reported on stack traces that were defined on applications and third-party libraries 
(the tool only analyzed the last version available).

\emph{\textbf{Exception Signaler Analysis.}}
This module is responsible for classifying each signaler according 
to its origin (i.e., Android Application Framework, Android Libcore, Application, Library). 
Table~\ref{tab:signalers} presents the heuristics adopted in this classification.
To conduct this classification, we provide this module with the information
comprising all Java packages that compose: the Android Platform;
 the Android Libcore; and each analyzed Application. To discover the packages for the first two origins
we can use the Android specification.
To discover the packages for the third origin, the application itself, this module
extracts the manifest files of each Android app, which defines the main packages that the applications consist of.
 If this file is not available, the tool recursively analyzes the 
structure of source-code directories composing the application, and filters out the cases in which the application also includes the source code of reused libraries.
Then, based on this information and using pattern matching 
between the signaler name and the packages, this module identifies 
the origin of the exception signalers.

The exceptions are considered to come
 from libraries if their packages are neither defined 
within the Android platform, nor on core libraries, nor on applications.
Table~\ref{tab:signalers} summarizes this signaler classification.

@@

After distilling the information available on the exception stack traces, we could find 
the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:topten} presents a list of the top 10 root exceptions found in the study - 
 ranked by the number of distinct projects on which they were reported. 
This table also shows how many times the signaler of such an exception was a method defined on
the Android platform, the Android libcore, the application itself or a third-party library -
 following the classification presented in Table  \ref{tab:signalers}.

@@

\finding{%
For native methods, even checked exceptions can be
thrown without being documented on the exception interface.
It violates best practices III and IV and represent a bug hazard
 hard to diagnose.
}

>>>>

\begin{abstract}
This paper reports two studies:  a repository mining study whose goal was to identify exception handling bug hazards in Android apps; and a survey-based study whose goal was to assess the perspectives of developers concerning such bug hazards. The repository mining study mined the exception stack traces embedded in 159,048 issues reported in Android projects hosted on GitHub (482 projs.)  and Google Code (157 projs.).  Overall 6,005 exception stack traces were extracted and analyzed in combination with source code and bytecode analysis. Some bug hazards were consistently detected, such as:  unexpected exception wrappings (e.g., errors being wrapped in checked exceptions) and NullPointerExceptions as the major cause of reported crashes -- reported in 28\% of all mined stack traces found in about 50\% of all analyzed projects. Finally, a survey was applied to some developers involved in some of the GitHub Android projects analyzed in our study and their perspective about the bug hazards and how they deal with them was assessed and discussed.

@@

The number of mobile apps is increasing at a daily rate. If on the one hand they
extend phones' capabilities far beyond the basic calls, on the other hand they have to 
cope with an increasing number of exceptional conditions 
(e.g., faults in underlying middleware or hardware; compatibility issues~\cite{McDon13}; 
memory and battery restrictions; noisy external resources~\cite{Zhang12}).

@@

However, such mechanisms are more often than not the least 
understood and tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling,yuan:2014.osdi}.
As a consequence they may inadvertently negatively affect the system: exception-related code may introduce failures such as 
uncaught exceptions~\cite{jo2004uncaught, Zhang12} -- 
which can lead to system crashes, making the system even less robust~\cite{coelho2011unveiling}.

In Java, when an application fails due to an uncaught exception, 
it automatically terminates, while the system prints a stack trace to the console, 
or to a log file~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
of the thrown exception and the ordered list of methods that were active on the call stack before 
the exception occurred~\cite{gosling2000java,bloch2008effective}.
When available, the exception stack traces provide a useful source of information about system crashes ~\cite{bettenburg2008makes} which 
can enable different kinds of post-mortem analysis and support:  debugging~\cite{schroter2010stack}, 
bug classification and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  
automated bug fixing~\cite{sinha2009fault} and fault-proneness prediction models~\cite{kim2013predicting}.

@@

This work is conducted in two phases. First, a mining study performs a post 
mortem analysis of the exception stack traces 
included in issues reported on Android projects hosted on GitHub and Google Code.
The goal of this study is to investigate whether the reported exception stack traces
 can reveal common \emph{bug hazards} in the exception-related code.  
A \emph{bug hazard}~\cite{binder2000testing}  is a circumstance that increases the 
chance of a bug being present in the software. An example of a bug hazard can 
be a characteristic of  the exception-related code which can increase the likelihood 
of introducing the aforementioned uncaught exceptions.
Second, we conducted an exploratory survey with the Android developers involved in the mined projects
to assess their perspective about the exception handling bug hazards found.

To guide this investigation we compiled general guidelines on how to use Java
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.
Then, using a custom tool called ExceptionMiner,
 which we developed specifically for this study, we mine stack traces from the issues reported 
in 482 Android projects hosted on GitHub and 157 projects hosted on Google Code.
Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
The exception stack trace analysis was augmented by means of bytecode 
and source code analysis for the exception-related code of the Android 
platform and Android applications. Some \emph{bug hazards} consistently detected 
during this mining study include:

@@

   \item  Cross-type exception wrappings, such as an OutOfMemoryError wrapped in a checked exception.
Trying to handle an instance of OutOfMemoryError ``hidden" in  a checked exception may bring the program
 to an unpredictable state. Such wrappings suggest that, when (mis)applied, the exception wrapping can make 
the exception-related code more complex and negatively impact the application robustness.

  \item  Undocumented runtime exceptions raised by the Android platform (35 methods) and  third-party libraries (44 methods) --
 which correspond to 4.4\% of the reported exception stack traces.
In the absence of the ``exception specification" of third-party code, it is difficult or 
even infeasible for the developer to protect the code against ``unforeseen'' exceptions. 
Since in such cases the client usually does not have access to the source code, such undocumented 
exceptions may remain uncaught and lead to system crashes.

@@

 \item A multitude of programming mistakes -- approximately 52\% of the reported stack traces can 
be attributed to programming mistakes. In particular, 27.71\% of all stack traces contained a java.lang.NullPointerException 
as their root cause.

@@

The exploratory survey was conducted with 70 Android developers involved in one or more of the 
GitHub Android projects whose issues were mined. This survey reveals that only few developers (3\% of respondents)
 knew about the undocumented checked exceptions signaled by native C code of the Android platform. 
Moreover, most of the developers recognized that cross-type exception wrappings may
negatively impact the application robustness. The uncaught exceptions due to errors in programming 
logic, e.g. the NullPointerException, were identified by most developers (68\%) as the first or second main cause 
of application crashes.

 The high prevalence of NullPointerExceptions found in the mining study and confirmed in the exploratory survey
 is in line with findings of earlier research~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}, 
as are the undocumented runtime exceptions signaled by the Android Platform~\cite{kechagia2014}.

Some of the findings of our mining study emphasize the impact of these bug hazards on the application robustness 
by mining a different source of information as the ones used in previous works. The present work 
mined issues created by developers on GitHub and Google Code, while  previous research analyzed crash reports and automated test reports.
 Furthermore, our work points to bug hazards that were not detected by previous research (i.e., cross-type wrappings, undocumented checked exceptions and undocumented runtime exceptions thrown by third-party libraries) which represent new threats to application robustness.
Moreover, we perform  the first exploratory survey study whose
goal was to assess developers' perspective regarding a set of exception handling bug hazards
as well as how developers deal with exceptions and prevent crashes while developing.

Our findings point to threats not only to the development of robust Android apps, 
but also to the development of any robust Java-based system. 
Hence, the study results are relevant to Android and Java developers who may underestimate the effect of such
 \emph{bug hazards} on the application robustness,
and who have to face the difficulty of preventing them.
Moreover, such \emph{bug hazards} 
call for improvements of languages (e.g. to prevent null pointer dereferences) and tools to better support exception handling in Android and Java environments.

The remainder of this paper is organized as follows. 
Section~\ref{sec:back} provides the necessary
background on the Android platform and the Java exception model. 
Section~\ref{sec:study} presents the study design. 
Section~\ref{sec:exceptionminer} describes the ExceptionMiner tool we developed to conduct our study.
Section~\ref{sec:result} reports the findings of our study.
Section~\ref{sec:disc} provides a discussion of the wider implications of our results,
 presents the threats to validity associated with the mining study and discusses the limitations
of the survey-based study. 
Finally Section~\ref{sec:rele} describes related work, 
and Section~\ref{sec:conc} concludes the paper and outlines
directions for future work.
\section{Background}
\label{sec:back}

@@

\emph{\textbf{Exception Types.}} In Java, exceptions are represented according to a class hierarchy, in which
 every exception is an instance of the Throwable class, and can be of three kinds: the checked exceptions
(extends Exception), the runtime exceptions (extends RuntimeException) and errors
(extends Error)~\cite{gosling2000java}. Checked exceptions received their name
 because they must be declared in the method's \emph{exception interface} (i.e., the list of exceptions that a method 
might raise during its execution) and the compiler statically checks if
 appropriate handlers are provided within the system.
Both runtime exceptions and errors are also known as ``unchecked exceptions'', as 
they do not need to be specified in the method \emph{exception interface} and do not trigger any 
compile time checking.

By convention, instances of Error represent  unrecoverable conditions which usually result
from failures detected by the Java Virtual Machine due to resource limitations, such as OutOfMemoryError.
Normally these cannot be handled inside the application.  Instances of RuntimeException are implicitly 
thrown by the Java runtime environment when a program violates 
the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
error, null pointer references). Some programming languages react to such errors by immediately terminating the program, while
other languages, such as C++, let the program  continue
 its execution in some situations such as the out-of-bounds array index.
According to the Java Specification~\cite{gosling2000java} programs are not 
expected to handle such runtime exceptions signaled by the runtime environment.

User-defined exceptions can be either checked
 or unchecked, by extending either Exception or RuntimeException. There is a long-lasting debate 
about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}.
%\footnote{152 questions in Stackoverflow are related to this debate}. 
Section~\ref{sec:best} presents
a set of best practices related to each of them.

@@

 A common way of  propagating exceptions in Java programs is through
 exception wrapping (also called chaining):
One exception 
is caught and wrapped in another one which is then thrown instead. Figure~\ref{fig:wrapping} shows 
an exception stack trace which illustrates such exception wrapping. 
For simplicity, in this paper we will refer to ``exception stack trace'' as just stack trace.
The bottom part of the stack trace is the \emph{root exception} (Figure~\ref{fig:wrapping}-A), which indicates
the first reason (root cause) for the exception thrown (in this case, the computer ran out of
memory). The top part of the stack trace indicates the location of the exception
manifestation, which we will refer to as the \emph{exception wrapper} in this paper (Figure~\ref{fig:wrapping}-C). The
execution flow  between the root exception and the wrapper may
include other intermediate exception wrappers (Figure~\ref{fig:wrapping}-D). At all levels, the exception
\emph{signaler} is the method that threw the exception, represented in the
stack trace as the first method call below the exception declaration (Figure~\ref{fig:wrapping}-B).

@@

Our study focused on the exception stack traces contained in issues of Android projects (hosted on GitHub and Google Code).
To support our investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
which extracts the exception stack traces embedded in issues 
and combines stack trace information with source code and bytecode 
analysis. Moreover, we use manual inspection to augment
 the understanding of stack traces and support further discussions and insights (Section~\ref{sec:manual}).
In this study we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively.

Figure~\ref{fig:overview} gives an overview of our study.
First,  the issues reported in Android projects hosted on GitHub (1) and Google Code (2) are recovered.
Then the stack traces embedded in each issue are extracted and distilled (3).
The stack trace information is then combined with source code and bytecode analysis in order to discover 
the type of the exceptions (5) reported in the stack traces (e.g., error, runtime, checked),
and the origin (6) of such exceptions 
(e.g., the application, a library, the Android platform). 
Manual inspection steps (4, 7, 9) are used to support the mining process  and 
the search for \emph{bug hazards}  (8). The next sections detail each step of this mining process.

Our study focuses on open-source apps, since the information needed to perform our study
cannot be retrieved from commercial apps, whose issue report systems and 
source code are generally not publicly available. 
Open source Android apps have also been the 
target of other research~\cite{Linar13,Ruiz12} addressing \emph{reuse} and API stability.

@@

\subsection{Android Apps on GitHub}
\label{sec:git}

@@

This study uses the dataset provided by the GHTorrent project~\cite{Gousi13}, 
an off-line mirror of the data  offered through the GitHub API.  
To identify Android projects, we performed a case insensitive search for the
term ``android'' in the repositories' names and short descriptions.  
Up to 23 February 2014,  when we queried GHTorrent, this resulted in 2,542 repositories. Running the ExceptionMiner tool 
 on this set we observed that 589 projects had at least one issue containing a stack trace.

Then we performed a further clean up, inspecting the site of every Android project
reporting at least one stack trace, to make sure that they represented real
mobile apps. During this clean up 107 apps were removed because they were either
example projects (i.e., toy projects) or tools to support Android development
(e.g. Selendroid, Roboeletric -- tools to support the testing of Android apps).
The filtered set consisted of 482 apps. This set of 482 projects contained a total of 31,592 issues from which 4,042 exception stack traces 
were extracted.

Issues on GitHub are different from issues in dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, GitHub uses a more open ended tagging system, where
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none or an arbitrary set of labels depending 
on its repository. Table~\ref{tab:lables} illustrates the occurrences of different labels 
on the issues including exception stack traces. Regardless of the issue labels, every exception stack
trace may contain relevant information concerning the exception structure of the
projects analyzed, and therefore can reveal \emph{bug hazards} in the exception-related code.
Because of this, we opted for not restricting the analysis to just defect issues.

@@

\subsection{Android Apps in Google Code}
Google Code contains widely used open-source Android apps (e.g. K9Mail\footnote{K9Mail moved to GitHub but as a way of not loosing the project history it advises their users to report bugs in the Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}).
However, differently from GitHub, Google Code does not provide an API to access the information related
 to hosted projects.\footnote{Google Code used to provide a Web service to its repositories, but this was deactivated in June 2013 in what Google called a ``clean-up action''.}
To overcome this limitation we needed to implement a Web Crawler (incorporated in the ExceptionMiner tool described next) that navigates 
 the web interface of Google Code projects extracting all issues and issue comments and storing them in a relational database for later analysis.
To identify Android projects on Google Code, we performed a similar heuristic: we performed a case insensitive search 
(on the Google Code search interface) for the term ``android''. In January 2014, when we queried Google Code, this resulted in a list of 788  projects. This list comprised the seeds sent to our Crawler.

The Crawler retrieved all issues and comments for these projects.
From this set, 724 projects defined at least 1 issue. Running the ExceptionMiner tool 
 on this set we observed that 183 projects had at least one issue containing an exception stack trace.
 Then we performed further clean up (similar to the one described previously) inspecting the site 
of each project. As a result we could identify 157 Android projects.  This set contained  127,456 issues in total,
 from which 1,963 exception stack traces were extracted. Table~\ref{tab:lables} illustrates the occurrences of different labels 
on the issues including exception stack traces. Differently from GitHub, on Google Code most of 
the issues were labeled as ``Defect''. However, based on the same assumption described for the GitHub repository
 we considered all issues reporting stack traces (regardless of their labels).

@@

\emph{\textbf{Repository Connectors.}} This component enables the connection 
with issue repositories. In this study two main connectors were created: one which connects to 
the GHTorrent database, and a Google Code connector which is comprised of a Web Crawler
 that can traverse the Google Code web interface and extract a project's issues. 
Project meta-data and the issues associated with each project are stored in a relational 
database.

\emph{\textbf{Exception Stack Trace Distiller.}}
This component combines a parser (based on regular expressions) 
and heuristics able to identify and filter exception names and stack traces inline with text. 
This component distills the information that composes a stack trace.
 Some of the attributes extracted from the stack trace are
 the root exception and its signaler, as well as the exception wrappers and their corresponding signalers. 
This component also distills fine grained information of each attribute such as the classes and packages associated with them.
In contrast to existing issue parsing solutions such as Infozilla, our parser
 can discover stack traces mixed with log file information\footnote{In several 
exception stack traces, the exception frames were preceded by logging information e.g., 
\texttt{03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread\-.java:127)} which could not be detected by existing tools.}.

\emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
stack traces every exception defined in a stack trace needs to be classified according to its type
(e.g. Error, checked Exception or RuntimeException). The module responsible for this analysis 
uses the Design Wizard framework \cite{Brunet09} to inspect the bytecode of the exceptions reported 
in stack traces.  It walks up the type hierarchy of a Java exception until it reaches 
a base exception type. Hence in this study the bytecode analysis was used to discover the type 
of each mined exception when the jar file of such an exception was available in the project or 
in a reused  Java library. A specific implementation (based on source code analysis)
 was needed to discover the exception type when the bytecode was not available.
With this module we analyzed all exceptions defined in the Android platform (Version 4.4, API level 19),
which includes all basic Java exceptions that can be thrown during the app execution,
and exceptions thrown by Android core libraries. Moreover, we also analyzed the 
exceptions reported in stack traces that were defined by applications and third-party libraries 
(the tool only analyzed the last version available).

\emph{\textbf{Exception Signaler Analysis.}}
This module is responsible for classifying each signaler according 
to its origin (i.e., Android Application Framework, Android Libcore, Application, Library). 
Table~\ref{tab:signalers} presents the heuristics adopted in this classification.
To conduct this classification, we provide this module with the information
comprising all Java packages that compose: the Android Platform;
 the Android Libcore; and each analyzed Application. To discover the packages for the first two origins
we can use the Android specification.
To discover the packages for the third origin, the application itself, this module
extracts the manifest files of each Android app, which defines the main packages that the applications consist of.
 If this file is not available, the tool recursively analyzes the 
structure of source code directories composing the application, and filters out the cases in which the application also includes the source code of reused libraries.
Then, based on this information and using pattern matching 
between the signaler name and the packages, this module identifies 
the origin of the exception signalers.

The exceptions are considered to come
 from libraries if their packages are neither defined 
within the Android platform, nor in core libraries, nor in the applications.
Table~\ref{tab:signalers} summarizes this signaler classification.

@@

After distilling the information available in the exception stack traces, we could find 
the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:topten} presents a list of the top 10 root exceptions found in the study -- 
 ranked by the number of distinct projects in which they were reported. 
This table also shows how many times the signaler of such an exception was a method defined by
the Android platform, the Android Libcore, the application itself or a third-party library --
 following the classification presented in Table  \ref{tab:signalers}.

@@

\finding{%
For native methods, even checked exceptions can be
thrown without being documented in the exception interface.
This violates best practices III and IV and represents a bug hazard
 hard to diagnose.
}
